\Title

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : C

Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task that is used to solve the ecological task described in section~\ref{sec:task}. Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics inspired by neuroscience are the main methodological differences with that previous study. First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) =  (21, 17, 17)$ and define as many classes as the number of motions (directions and speeds): $N_\class = XXX8 \times XXX3$. To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels. Furthermore, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}. Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we thus apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration). This is consistent with neuroscientific principles because, due to the limits of the conduction delay along horizontal connections, the synaptic delay is related, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one. Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities arrangements. This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable. Since this arrangement of polarities is independent of speed, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel. This is similar to the computation done for complex cells in primary visual cortex.

\begin{figure}%[h!]%[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/HDSNN_conv.pdf}
    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
    \caption{
    % \textbf{- HD-SNN illustration. From left to right:} A 3D representation of the event stream as output of the DVS. A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis. There are as many kernels as the possible number of motion classes in the recording. For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid. For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e. softmax function). The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike. With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class. }
    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time). A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels. The convolution is summed over the two polarities. If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution. We restrict the illustration to a 2D representation and to 2 possible classes (green and orange) that are associated with different motion directions. \textit{(Middle)}~For each position (address, time), one can compute the activation resulting from the convolution. The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function). The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel). \textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class. Note that the position of the output spikes does not systematically correspond to the position of the input spikes but only when enough evidence is obtained.}
    \label{fig:model}
\end{figure}

\subsection{Supervised learning of the motion detection task}
%
Since the model is fully differentiable, we can now implement a supervised learning rule. This rule was implemented using the binary input events as inputs and the corresponding motion direction labels as the desired output. The loss function of the MLR model is the binary cross entropy at the output of the classification layer. The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions. Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images. However, the supervised optimization of this MLR model will involve adjusting the weights of the kernels. As a result, the error is only propagated back to the spatial locations of these most active cells. This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.

Finally, the output of the MLR model is an event-based representation that predicts the probability of each motion at each position and time. Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task. We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities. Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at this time. This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted). These calculations are performed on a different input dataset than the one used in the training or validation steps. The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN} (see Data Availability). %

>>>>

Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task used to solve the ecological task described in section~\ref{sec:task}. Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study. First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) = (21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = XXX8 \times XXX3$. To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels. We also included a prior in the selectable motions, as there is a prior for slow speeds in natural scenes~\citep{vacher_bayesian_2018}. Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration). This is consistent with neuroscientific principles because, due to the limits of conduction delay along horizontal connections, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons. Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities. This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable from a standard one. Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel. This is similar to the computation done for complex cells in primary visual cortex.

\begin{figure}%[h!]%[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/HDSNN_conv.pdf}
    \caption{
        \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot a 2D representation of the input event stream as a raster plot (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time). A spatio-temporal convolution is applied to the dense representation of the input with $2$ different convolution kernels (the green and orange kernels), which will define the output channels. The convolution is summed over the two polarities. Since we have two axes $X$ and $Y$ to represent the presynaptic addresses, like the pixel grid of a DVS, this results in a 3D convolution. Here we simplify the illustration to a 2D representation and to 2 possible classes (green and orange) associated with different directions of motion. \textit{(Middle)}~For each position (address, time) one can compute the activation resulting from the convolution. The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function). The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel). \textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated with thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated with a particular motion class. Note that the position of the output spikes does not systematically correspond to the position of the input spikes, but only when enough evidence is obtained.
    }
    \label{fig:model}
\end{figure}

\subsection{Supervised learning of the motion detection task}
%
Since the model is fully differentiable, we can now implement a supervised learning rule. This rule was implemented using the binary input events as inputs and the corresponding motion's labels as the desired output. The loss function of the MLR model is the binary cross entropy at the output of the classification layer. The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions. Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images. However, the supervised optimization of this MLR model will adjust the weights of the kernels. As a result, the error is only propagated back to the spatial locations of these most active cells. This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies, each of size $TTT \times XXX \times YYY$ and a learning rate of $10^{-5}$). We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.

Finally, the output of the MLR model is an event-based representation that predicts the probability of each motion at each position and time. Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task. We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities. Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at that time. This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted). These calculations are performed on a different input data set than the one used in the training or validation steps. The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN} (see Data Availability). %
